The Iterative Signature Algorithm for Gene Expression Data

### 2010-03-15 eisa/inst/doc/EISA_tutorial.Rnw
Update EISA tutorial.

The easiest way to create a heatmap of a single module is to call the
\Rfunction{ISA2heatmap} function. You need to specify which module you
want to plot and also the \Rclass{ExpressionSet} object that is being
analyzed. Note that by default ISA normalizes the expression data
before running the module detection, but \Rfunction{ISA2heatmap} plots
the raw values by default, optionally scaled. 
If you want to plot the normalized values, then you need supply the
\Rargument{norm} argument to \Rfunction{ISA2heatmap}, see the manual
for details. The result is in Fig.~\ref{fig:heatmap}.
<<heatmap,results=hide,include=FALSE,eval=FALSE>>=
ISA2heatmap(modules, 1, ALL, norm="sample", scale="none")
@

\begin{figure}
\centering
<<heatmap-real,fig=TRUE,echo=FALSE,results=hide>>=
<<heatmap>>
@
\caption{Heatmap of the first ISA module, plotted via the
\Rfunction{ISA2heatmap} function. The expression values are
sample-wised normalized.}
\label{fig:heatmap}
\end{figure}

@@

You can also use the \Rpackage{biclust} package to create image
plots, by calling the \Rfunction{drawHeatmap} function. The result is
in Fig.~\ref{fig:drawHeatmap}.
<<heatmap biclust pre,echo=FALSE,results=hide>>=
if (require(Cairo)) {
  CairoPNG("plot-heatmap.png", 400, 400 )  
} else {
  png("plot-heatmap.png", 400, 400 )
}
@ 
<<heatmap>>=
drawHeatmap(exprs(ALL[featureNames(modules),]), Bc, number=1) 
@
<<heatmap post,echo=FALSE,results=hide>>=
dev.off() 
@ 
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{plot-heatmap.png}
  \caption{Heatmap of the first transcription module, plotted with the
  \Rfunction{drawHeatmap} function from the \Rpackage{biclust}
  package.}
  \label{fig:drawHeatmap}
\end{figure}

@@

The \Rpackage{ExpressionView} package~\cite{ev} tries to reorder the
rows and columns of the expression matrix in such a way that the
genes/samples that appear together in a module, are placed right
beside/above each other. The package directly supports the
\Rclass{ISAModules} and \Rclass{ExpressionSet} classes, see its
documentation for details.

>>>>

The easiest way to create a heatmap of a single module is to call the
\Rfunction{ISA2heatmap} function. You need to specify which module you
want to plot and also the \Rclass{ExpressionSet} object that is being
analyzed. Note that by default ISA normalizes the expression data
before running the module detection, but \Rfunction{ISA2heatmap} plots
the raw values by default, optionally scaled. 
If you want to plot the normalized values, then you need supply the
\Rargument{norm} argument to \Rfunction{ISA2heatmap}, see the manual
for details. 
<<heatmap,results=hide,eval=FALSE>>=
ISA2heatmap(modules, 1, ALL, norm="sample", scale="none", 
            col=heat.colors(12))
@

We also create a color bar.
<<heatmap-cb,results=hide,eval=FALSE>>=
sc <- function(x, ab) {
  x <- x-min(x)
  r <- range(x) 
  x <- x/(r[2]-r[1])
  x * (ab[2]-ab[1]) + ab[1]
}
cb <- sc(1:12, range(exprs(ALL)[getFeatureNames(modules, 1)[[1]],
                                      getSampleNames(modules, 2)[[1]] ]))
par(mar=c(3,4,5,2)+0.1)
image(rbind(cb), axes=FALSE)
axis(2, at=sc(1:12, 0:1), cb, label=format(cb, digits=2))
@

The result is in Fig.~\ref{fig:heatmap}.

<<save heatmap files,echo=FALSE,print=FALSE,results=hide>>=
png("plot-heatmap.png", 400, 400)
<<heatmap>>
dev.off()
png("plot-heatmap-colorbar.png", 100, 400)
<<heatmap-cb>>
dev.off()
@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{plot-heatmap.png}%
\includegraphics[width=0.2\textwidth]{plot-heatmap-colorbar.png}
\caption{Heatmap of the first ISA module, plotted via the
\Rfunction{ISA2heatmap} function.}
\label{fig:heatmap}
\end{figure}

@@

@@

The \Rpackage{ExpressionView} package tries to reorder the
rows and columns of the expression matrix in such a way that the
genes/samples that appear together in a module, are placed right
beside/above each other. The package directly supports the
\Rclass{ISAModules} and \Rclass{ExpressionSet} classes, see its
documentation for details.
